{Investigating the Impact of Macroeconomic Indicators on the FTSE 250 Index Using Statistical and Machine Learning Methods}

### 2024-08-26 tex/paper.tex
revised writeup

\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{tikz}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[numbers]{natbib}
\usepackage{natbib}
\usepackage{parskip}
\usepackage{float}
\usepackage{tablefootnote}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\providecommand{\keywordname}{\textbf{Keywords:}} % Customize how "Keywords:" appears
\newcommand{\keywords}[1]{%
  \par\addvspace{\baselineskip}% Adds some space above the keywords
  \noindent\keywordname\enspace\ignorespaces#1 % Displays "Keywords:" followed by the actual keywords
}
\newcommand{\citeboth}[1]{\citeauthor{#1} \citep{#1}}

@@

\title{Investigating the Impact of Macroeconomic Indicators

on the FTSE 250 Index Using Statistical and Machine Learning Methods}
\author{Mateen Khan}
\date{\today}

\begin{document}

@@

The remainder of this paper will be organised as follows. 
Section \ref{sec:lit} provides a brief review of the literature in this problem space and explains the rationale behind the selection of macroeconomic variables.
Section \ref{sec:meth} presents the methodology. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
the study.

@@

The study uses monthly FTSE250 price data between October 1992 and March 2024,
thus providing 379 observations. The data for the macroeconomic indicators 
was also captured on a monthly basis in the same time frame, except for the 
interest rate. This is because the Bank of England's Monetary Policy 
Committee, 
which is responsible for setting interest rates, meets on an \textit{ad hoc} 
basis\footnote{The MPC typically meets roughly every six weeks, however, in uncertain economic 
climes it may convene far more frequently.}
to decide on whether the interest rate should be changed. The sources 
are presented in Table \ref{table: tabl}.

\clearpage
\begin{table}[h!]
    \centering
    \caption{Data sources}
    \label{table: tabl}
    \begin{tabular}{llll}
        \toprule
        \textbf{Variable} & \textbf{Frequency} & \textbf{Notation} & \textbf{Source} \\
        \midrule
        FTSE 250 Index & Monthly & FTSE$_{250}$ & Yahoo Finance \\
        Consumer Price Index & Monthly & CPI & Office for National Statistics (ONS) \\
        USD/GBP Exchange Rate (£) & Monthly & EXCHG & Federal Reserve of St Louis \\
        Interest Rate ($\%$) & Variable & INT & Bank of England \\
        Money Supply (M3) (millions £) & Monthly & M3 & Bank of England \\
        \bottomrule
    \end{tabular}
\end{table}

@@

To address the variability in the interest data, Algorithm \ref{alg:interest_rate_adjustment}
was used to convert the data into a monthly format, ensuring a uniform
frequency across the variables.

@@

The result is a dataset $X$ composed of monthly values for the FTSE250
price, CPI value, USD/GBP exchange rate, interest rate, and M3 money supply.

@@

The CPI, with an average of 88.36, has 
experienced significant fluctuations, highlighting periods of high inflationary pressure such as the onset of the Russia-Ukraine conflict in 2022, and periods of stable, low
inflation such as the early to late 2010s.

Interest rates, have varied widely from $0.1\%$ to $8.4\%$, with a standard deviation of $2.52\%$, reflecting the 
increasing role played by monetary policy in the UK since May 1997 when operational responsibility for setting interest rates was tranferred from the Chancellor of the Exchequer to the Bank of England (\citeboth{king1997changes}). The Bank of England was forced to adjust interest rates 
several times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 financial crisis to stimulate the economy and the gradual rate hikes in response to 
post-COVID inflationary pressures.

The M3 money supply has shown substantial variability, ranging from 494,181 to 3,704,873, 
in large part influenced by the various Quantitative Easing (QE) programs implemented by the Bank of England 
in response to the recessions onset by the 2008 financial crisis and the COVID-19 pandemic. These programs aimed to inject liquidity into the economy to stimulate economic activity, thus inflating the overall money supply.

@@

The study uses an autoregressive distributed lag (ARDL) model in conjunction with 
an error correction model (ECM) to examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250
and to quantify these relationships. Prior to constructing the model a natural logarithmic transform was applied
on the variables. This was done so that the coefficients can be interpreted easily as percentage changes.

@@

The ARDL model specifies the dependent variable as a function of its own 
lagged values and the lagged values of independent variables. 
The variables are permitted to be of differing stationarity. The stationarity 
of the variables was tested using the Augmented Dickey-Fuller 
Unit Root Test found within the \texttt{tseries} package in R. The optimal lags for each of the 
variables were calculated using the Schwarz Information Criterion, contained
within the auto\_ardl function in the ARDL package in R.

@@

The intercept term was given by $\gamma_0$. The 
remaining $\gamma_i, \delta_i$ coefficients represented the short-run 
relationships between the variables. The $\lambda$ coefficient 
described the speed at which short-run deviations from the long-run equilibrium
were adjusted. A negative $\gamma$ implies that there existed a correction mechanism that guided the deviation 
back to the long-run equilibrium. These coefficients, as well 
as the error correction term ECT, were computed using the \texttt{recm} function in the 
\texttt{ARDL} library in R.

The \texttt{{bounds\_f\_test}} in 
the \texttt{ARDL} library, based on the bounds test approach 
outlined in \citeboth{pesaran2001}, was used to calculate the 
F-statistic of the ARDL model and subsequently test for cointegration.

@@

To confirm that the model did not display autocorrelation, heteroskedasticity, non-normality, or structural instability, the Box-Ljung, Breusch-Pagan, Shapiro-Wilk, and CUSUM tests were conducted, respectively

@@

Neural networks work best with data that is specified between a specific 
range (\citeboth{mendelsohn1993}).
Also, based on the unit root tests in Table \ref{table: unitroot} the variables in the dataset $X$, described in \ref{section: prep},
only achieve stationary at first difference.
Therefore, all of the 
variables in the dataset $X$ were normalised using a z-score normalisation and differenced,
as described in Algorithm \ref{alg:z_score_differencing}.

@@

The inputs into the network were 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\), from the differenced, normalised dataset $Z$

@@

The output was a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\) from the dataset $Z$, representing differenced monthly values in the FTSE250.

@@

The radial basis function neural network was programmed into Python. The Bayesian kernel regression approach outlined by 
\citeboth{bishop1995}, using the GaussianProcessRegressor package in Python was employed to model the network. For a 
detailed description of this construction please refer to the \hyperref[sec: appendix]{Appendix}.

The approach outlined in the \hyperref[sec: appendix]{Appendix} was used to conduct the train-test process:
\begin{enumerate}
    \item The optimal hyperparameters in the covariance function \eqref{eq: gimq} 
    were computed using GridSearch Cross Validation. These hyperparameters were:
    \begin{enumerate}
        \item $\alpha$: a regularisation parameter representing a small quantity
        that was added to the diagonals of the covariance matrix to ensure it was 
        invertible. 
        \item $\beta$: the power in the covariance function, which controlled the shape of the function.
        \item $\sigma$: a parameter which controlled how much of the noise in the data the kernel accounted for. It moderated the risk of overfitting and underfitting.
        \item Length scale: a parameter that controlled the influence that each kernel was to have. Analogous to 
        the width that each radial basis function neuron possesses.
    \end{enumerate}
    \item Training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
    and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) were successively passed through the 
    covariance function $(x,x')$, to obtain the matrix of the pairwise covariance of 
    all training observations, $K$.
    \item A $4$-dimensional input space $X$ took in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
    calculated the posterior weights distribution. 
    Each possible weight vector $w$ represented a different function, explicitly 
    based on the radial basis function covariance function, that could fit the data. In the 
    hidden layer the expectation of this distribution was then taken, which
    combined all the weights and found the mean, weighted by the 
    respective probabilities. This represented the best functional estimate of the price movement. Based on this estimate 
    the predicted output was generated.
    \item The testing process was conducted similarly. The 
    posterior weight distribution that contained the information of the training phase became the new 
    prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
    were calculated, and then the predictive output was found (\citeboth{rasmussen2006}).

@@

\begin{table}[h!]
    \centering
    \caption{ARDL Bounds Test Critical Values}
    \begin{tabular}{lcccc}
        \toprule
        \textbf{Order} & \textbf{10$\%$} & \textbf{5$\%$} & \textbf{2.5$\%$} & \textbf{1$\%$} \\
        \midrule
        I(0) & $2.20$ & $2.56$ & $2.88$ & $3.29$ \\
        I(1) & $3.09$ & $3.49$ & $3.87$  & $4.37$ \\
        \bottomrule
    \end{tabular}
\end{table}

@@

The interest rate had a significantly
negative short-term impact on the FTSE250 price in the short term. 
This can be attributed to the immediate effect of interest rate hikes, 
which decrease the present value of future earnings according to the 
Discounted Cash Flow (DCF) valuation model, thereby 
affecting investor sentiment and depressing stock prices. 
This finding was consistent with the studies by 
\citeboth{alam2009}, \citeboth{demir2019}, and \citeboth{neifar2023}.

The money supply had a significantly negative short-term effect on the 
FTSE250 price, as noted by \citeboth{olawale2014}. 
According to \citeboth{sellin2001}, this aligned with the Keynesian argument 
that an increase in money supply leads to expectations of a future 
monetary contraction. This expectation increases the demand for 
funds, driving up current interest rates and reducing stock prices 
through the same mechanism as described previously.

In the short-term, before structural shifts can materialise, the
weakening of the Pound against the Dollar had a positive, but 
insignificant effect. As the USD has been 
the global reserve currency for many decades, the depreciation of the Pound 
against it would lead to increased revenue in Pound terms for FTSE250 
companies with substantial overseas operations. 
This is because their foreign earnings, 
when converted back into Pounds, will be worth more, 
potentially increasing profits. However, the insignificant impact suggests 
that FTSE250 firms are resilient to short-term fluctuations in the stock market 
price. The possible reasons for this are that they have diversified 
operations, meaning fluctuations against the Dollar will not threaten 
their profits and they use hedging strategies involving financial instruments 
such as currency forwards which protect against volatility.

The insignificance of inflation, defined as a positive 
increase in the CPI, confirmed the findings of 
\citeboth{gultekin1983}, \citeboth{firth1979} and \citeboth{neifar2023}. 
The CPI as a measure does not uniformly represent all sectors of the economy, 
and may consequently be unrepresentative of inflationary changes in the 
sectors that predominate the FTSE250. In addition, different sectors 
are more vulnerable to inflation than others which may distort the effect 
of inflation even further. Moreover, inflation may be anticipated 
by investors and already be factored into stock prices.

The error correction term 
was negative and statistically significant ensuring that long-run equilibrium 
can be achieved. The adjustment speed of $0.1794$ indicated that 
roughly $18\%$ of the fluctuation in the previous month's price caused by the 
macroeconomic variable movements was adjusted back to the long-run equilibrium in the subsequent month. This implies 
that it takes approximately 6 months for the FTSE250 to return to its long-run equilibrium following 
changes in the selected macroeconomic variables.

@@

The CPI did not have a significant effect on the FTSE250 price over the 
long-run, lending support to the findings of 
\citeboth{gultekin1983} and \citeboth{firth1979}
that the Fisher hypothesis is valid for the UK. One of the 
reasons for this may be that, since stock prices represent the 
discounted value of future cash flows (which are real), 
they are not directly impacted by inflation in the long run. Moreover, 
firms are able to pass on nominal costs to consumers which reduces
the impact of inflation.
The freer role of the Bank of England in setting monetary policy could be 
another factor in this, as a stricter monetary policy would keep inflation
in check and obscure any long-term equilibrium relationship.

@@

The USD/GBP exchange rate significantly positively impacted the FTSE250 price, confirming the findings of 
\citeboth{wong2022}. That there is a significant relationship is not surprising, because whilst the FTSE250 is a more domestic indicator of the London Stock Exchange than the FTSE 100 Index, 
at times nearly $50\%$ of its earnings have come from abroad, indicating that 
the value of the Pound has a significant impact on this stock market. There are many reasons for why this effect is observed.
In the long run, the appreciation of the Pound would decrease the 
cost of imports, thereby increasing the 
profits of FTSE250 companies and inflating their share prices.
Moreover, a stronger Pound makes UK assets more appealing to foreign investors, as their investments would be worth more when converted back to their domestic currencies.
This would drive up demand for UK stocks and positively influence the 
FTSE250 price.

The interest rate was also found to have significant negative long-term and short-term effects on the stock market, confirming the findings of \citeboth{alam2009}, \citeboth{demir2019} and \citeboth{neifar2023}
This can be explained by the fact that higher interest 
rates increase the cost of borrowing for companies, leading to reduced investment 
and lower future earnings potential, in turn depressing stock prices. 
Furthermore, as per the flight-to-quality phenomenon, persistently higher interest rates make bonds and other fixed-income 
investments more attractive relative to stocks, leading investors to shift 
their portfolios away from equities and toward safer assets over time (\citeboth{asgharian2016}). 
In the short term, interest rate hikes immediately decrease the present value of earnings, as per the 
DCF valuation model, and depress prices.

@@

The short-term prediction yielded MAE = $0.07$ and 
$R^2 = 61\%$ for a 5-month time horizon. The results obtained are presented in Figure \ref{fig:shortmonthly}.
The significantly better performance of the model in the short-run
indicates that the selected macroeconomic variables possess
a stronger predictive power over shorter periods. This was highlighted 
earlier in the ARDL model, with the interest rate and money supply significantly
impacting the FTSE250 price in the short-run.

@@

\subsubsection{Long-term Forecast}
The model returned a mean absolute error of MAE $=0.16$ and 
$R^2 = 9\%$ for the long-term prediction, providing a 
$2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model. The results are presented 
in Figure \ref{fig:longmonthly}.

@@

Stock markets provide important information on the economy, 
and economic policymakers are vigilant in tracking the fluctuations 
in these markets, and try to forecast the 
forthcoming fluctuations based on policy changes that affect the macroeconomy.
Within this context, the study examined the 
impacts of the interest rate, USD/GBP
exchange rate, the Consumer Price Index, and 
M3 money supply market activity on
the FTSE250 over the October 1993 – April 2024 period.
The results suggested a cointegrated long-run
relationship between the macroeconomic variables and the stock market index. 
The long-run coefficients implied that
increases in the interest rate and depreciation in the exchange rate
significantly depress the FTSE250 value. 
Policymakers dealing with forecasting the UK stock markets
should consider the factors in the study, and to raise the stock
market's value, a lower interest rate regime should be adopted. This will 
stimulate consumer activity, reduce operating costs for firms, and make 
stocks more attractive investments, thereby increasing the stock market value. 
In addition, the Pound should be increased in value to lower import costs 
and increase the real value of firm earnings, both of which should drive up 
the value of the stock market. The combined effect of these changes will 
create a more stimulating business environment by increasing the capital 
available to firms for research and development, and more efficient 
capital and labour, in turn 
driving economic growth and prosperity. However, 
as with all economic policy shifts, these changes may have 
unintended consequences on other areas of the economy, potentially 
undermining any positive changes. Any policy changes must cautiously be made 
with this at the forefront.

@@

\section{Appendix}
\label{sec: appendix}

The GaussianProcessRegressor function in Python's Scikit library is based on the 
standard function

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{tikz}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[numbers]{natbib}
\usepackage{natbib}
\usepackage{parskip}
\usepackage{titlesec}
\usepackage{float}
\usepackage{tablefootnote}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\providecommand{\keywordname}{\textbf{Keywords:}} % Customize how "Keywords:" appears
\newcommand{\keywords}[1]{%
  \par\addvspace{\baselineskip}% Adds some space above the keywords
  \noindent\keywordname\enspace\ignorespaces#1 % Displays "Keywords:" followed by the actual keywords
}
\newcommand{\citeboth}[1]{\citeauthor{#1} \citep{#1}}

@@

\titleformat{\title}[block]{\normalfont\bfseries}{\thetitle}{1em}{}

\title{{Investigating the Impact of Macroeconomic Indicators
on the FTSE 250 Index Using Statistical and Machine Learning Methods}}
\author{
  \textbf{Mateen Khan} \\
  Mathematics Institute, University of Warwick  \\ 
  \\
  \textbf{Supervisor: Dr. Randa Herzallah} \\
  Mathematics Institute, University of Warwick
}
\begin{document}

@@

The remainder of this paper will be organised as follows. 
Section \ref{sec:lit} provides a brief review of the literature in this problem space and explains the rationale behind the selection of macroeconomic variables.
Section \ref{sec:meth} presents the methodology of this study. Section \ref{sec: results} discusses the results and Section \ref{sec: conc} concludes
the study.

@@

This study used monthly FTSE250 price data between October 1992 and March 2024,
thus providing 379 observations. The data for the macroeconomic indicators 
was also captured on a monthly basis in the same time frame, except for the 
interest rate. This is due to the
\textit{ad hoc} 
basis\footnote{The MPC typically meets roughly every six weeks, however, in uncertain economic 
conditions it may convene far more frequently.} on which the
Bank of England's Monetary Policy 
Committee decides to set interest rates. The sources 
are presented in Table \ref{table: tabl}.

\begin{table}[h!]
    \centering
    \caption{Data sources}
    \label{table: tabl}
    \begin{tabular}{llll}
        \toprule
        \textbf{Variable} & \textbf{Frequency} & \textbf{Notation} & \textbf{Source} \\
        \midrule
        FTSE 250 Index & Monthly & FTSE$_{250}$ & Yahoo Finance \\
        Consumer Price Index & Monthly & CPI & Office for National Statistics (ONS) \\
        USD/GBP Exchange Rate (£) & Monthly & EXCHG & Federal Reserve of St Louis \\
        Interest Rate ($\%$) & Variable & INT & Bank of England \\
        Money Supply (M3) (millions £) & Monthly & M3 & Bank of England \\
        \bottomrule
    \end{tabular}
\end{table}

@@

To address the variable frequency of the interest rate data, Algorithm \ref{alg:interest_rate_adjustment}
was used to convert the data into a monthly format.

@@

The result was a dataset $X$ composed of monthly values for the FTSE250
price, CPI value, USD/GBP exchange rate, interest rate, and M3 money supply.

@@

The CPI, with an average of 88.36, has 
experienced significant fluctuations, highlighting periods of high inflationary pressure such as the onset of the Russia-Ukraine conflict in 2022, and periods of stable, low
inflation such as the period between the early to late 2010s.

Interest rates have ranged from $0.1\%$ to $8.4\%$ with a standard deviation of $2.52\%$, reflecting the 
increasing role played by monetary policy in the UK since May 1997. This was when operational responsibility for setting interest rates was tranferred from the Chancellor of the Exchequer to the Bank of England (\citeboth{king1997changes}). The Bank of England was forced to adjust interest rates 
several times in response to worsening economic conditions, examples being the steep rate cuts during the 2008 financial crisis to stimulate the economy and the gradual rate hikes in response to 
post-COVID inflationary pressures.

The M3 money supply has shown substantial variability, ranging from 494,181 to 3,704,873, 
in large part influenced by the various Quantitative Easing (QE) programs implemented by the Bank of England 
in response to the recessions caused by the 2008 financial crisis and the COVID-19 pandemic. These programs aimed to inject liquidity into the economy to stimulate economic activity, thus inflating the overall money supply.

@@

The study used an autoregressive distributed lag (ARDL) model in conjunction with 
an error correction model (ECM) to examine the short-term and long-term effects of macroeconomic indicators on the price of the FTSE250
and to quantify these relationships. Prior to constructing the model a natural logarithmic transform was applied
on the variables. This was done so that the coefficients can be interpreted easily as percentage changes.

@@

The ARDL model specifies the dependent variable as a function of its own 
lagged values and the lagged values of independent variables. 
In this model the variables are permitted to be of differing stationarity. The stationarity 
of the variables was tested using the Augmented Dickey-Fuller 
Unit Root Test found within the \texttt{tseries} package in R. The optimal lags for each of the 
variables were calculated using the Schwarz Information Criterion, contained
within the \texttt{auto\_ardl} function in the \texttt{ARDL} library in R.

@@

The intercept term was given by $\gamma_0$. The 
remaining $\gamma_i, \delta_i$ coefficients represented the short-run 
relationships between the variables. The $\lambda$ coefficient 
described the speed at which short-run deviations from the long-run equilibrium
were adjusted. A negative $\gamma$ implied that there existed a correction mechanism that guided the deviation 
back to the long-run equilibrium. These coefficients, as well 
as the error correction term ECT, were computed using the \texttt{recm} function in the 
\texttt{ARDL} library in R.

The \texttt{{bounds\_f\_test}} function in 
the \texttt{ARDL} library, based on the bounds test approach 
outlined in \citeboth{pesaran2001}, was used to calculate the 
F-statistic of the ARDL model and then test for cointegration.

@@

To confirm that the model did not display autocorrelation, heteroskedasticity, {non-normality}, or structural instability, the Box-Ljung, Breusch-Pagan, Shapiro-Wilk, and CUSUM tests were conducted, respectively

@@

Neural networks work best with data that is specified between a specific 
range (\citeboth{mendelsohn1993}).
Moreover, based on the unit root tests in Table \ref{table: unitroot} the variables in the dataset $X$, described in \ref{section: prep},
only achieve stationary at first difference.
Therefore, all of the 
variables in the dataset $X$ were normalised using a z-score normalisation and differenced
as described in Algorithm \ref{alg:z_score_differencing}.

@@

The inputs into the network were 4-dimensional input vectors, \(\boldsymbol{\mathrm{x_i}}\), from the differenced and normalised dataset $Z$

@@

The output was a one-dimensional output vector, \(\boldsymbol{\mathrm{y_i}}\) from the dataset $Z$, representing differenced, normalised monthly values in the FTSE250.

@@

The radial basis function neural network was programmed in Python. The Bayesian kernel regression approach outlined by 
\citeboth{bishop1995}, using the \texttt{GaussianProcessRegressor} package in Python was employed to model the network\footnote{See \hyperref[sec: appendix]{Appendix}}.

This approach was then used to conduct the train-test process as follows:
\begin{enumerate}
    \item The optimal hyperparameters in the covariance function \eqref{eq: gimq} 
    were computed using GridSearch Cross Validation. These hyperparameters were:
    \begin{enumerate}
        \item $\alpha$: a regularisation parameter representing a small quantity
        that was added to the diagonals of the covariance matrix to ensure it was 
        invertible. 
        \item $\beta$: the power in the covariance function, which controlled the shape of the function.
        \item $\sigma$: a parameter which controlled how much of the noise in the data the kernel accounted for. It moderated the risk of overfitting and underfitting.
        \item Length scale: a parameter that controlled the influence that each kernel was to have. Analogous to 
        the width that each radial basis function neuron possesses.
    \end{enumerate}
    \item Training inputs $\boldsymbol{\mathrm{x}}_1, \boldsymbol{\mathrm{x}}_2,\ldots \boldsymbol{\mathrm{x}}_{n}$
    and corresponding outputs \( y = \{y_1, y_2, \dots, y_n\} \) were successively passed through the 
    covariance function $(x,x')$, to obtain the matrix of the pairwise covariance of 
    all training observations, $K$.
    \item A $4$-dimensional input space $X$ took in an individual training observation $(\boldsymbol{\mathrm{x}}_{i}, y_i)$ and 
    calculated the posterior weights distribution. 
    Each possible weight vector $w$ represented a different function, explicitly 
    based on the radial basis function covariance function, that could fit the data. In the 
    hidden layer the expectation of this distribution was then taken, which
    combined all the weights and found the mean, weighted by the 
    respective probabilities. This represented the best functional estimate of the price movement. Based on this estimate 
    the predicted output was generated.
    \item The testing process was conducted similarly. The 
    posterior weight distribution that contained the information of the training phase became the new 
    prior distribution. From this the posterior weight distributions for the test observations \eqref{eq: testw}
    were calculated, and then the predictive output was found by taking the expectation (\citeboth{rasmussen2006}).

@@

\begin{table}[h!]
    \centering
    \caption{ARDL Bounds Test Critical Values}
    \begin{tabular}{lcccc}
        \toprule
        \textbf{Order} & \textbf{10$\%$} & \textbf{5$\%$} & \textbf{2.5$\%$} & \textbf{1$\%$} \\
        \midrule
        I(0) & $2.20$ & $2.56$ & $2.88$ & $3.29$ \\
        I(1) & $3.09$ & $3.49$ & $3.87$  & $4.37$ \\
        \bottomrule
    \end{tabular}
\end{table}
\clearpage

@@

The interest rate has a significantly
negative short-term impact on the FTSE250 price in the short term. 
This can be attributed to the immediate effect of interest rate hikes, 
which decrease the present value of future earnings according to the 
Discounted Cash Flow (DCF) valuation model, thereby 
affecting investor sentiment and depressing stock prices. 
This finding was consistent with the studies by 
\citeboth{alam2009}, \citeboth{demir2019}, and \citeboth{neifar2023}.

The money supply has a significantly negative short-term effect on the 
FTSE250 price, as noted by \citeboth{olawale2014}. 
According to \citeboth{sellin2001}, this aligned with the Keynesian argument 
that an increase in money supply leads to expectations of a future 
monetary contraction. This expectation increases the demand for 
funds, driving up current interest rates and reducing stock prices 
through the same mechanism as described previously.

In the short-term, before structural shifts can materialise, the
weakening of the Pound against the Dollar has a positive, but 
insignificant effect. As the USD has been 
the global reserve currency for many decades, the depreciation of the Pound 
against it would lead to increased revenue in Pound terms for FTSE250 
companies with substantial overseas operations. 
This is because their foreign earnings, 
when converted back into Pounds, will be worth more, 
potentially increasing profits. However, the insignificant impact suggests 
that FTSE250 firms are resilient to short-term fluctuations in the stock market 
price. The possible reasons for this are that they have diversified 
operations, meaning fluctuations against the Dollar will not threaten 
their profits and they use hedging strategies involving financial instruments 
such as currency forwards which protect against volatility.

The insignificance of inflation, defined as a positive 
increase in the CPI, confirms the findings of 
\citeboth{gultekin1983}, \citeboth{firth1979} and \citeboth{neifar2023}. 
The CPI as a measure does not uniformly represent all sectors of the economy, 
and may consequently be unrepresentative of inflationary changes in the 
sectors that predominate the FTSE250. In addition, different sectors 
are more vulnerable to inflation than others which may distort the effect 
of inflation even further. Moreover, inflation may be anticipated 
by investors and already be factored into stock prices.

The error correction term 
is negative and statistically significant ensuring that long-run equilibrium 
can be achieved. The adjustment speed of $0.1794$ indicates that 
roughly $18\%$ of the fluctuation in the previous month's price caused by the 
macroeconomic variable movements was adjusted back to the long-run equilibrium in the subsequent month. This implies 
that it takes approximately 6 months for the FTSE250 to return to its long-run equilibrium following 
changes in the selected macroeconomic variables.

@@

The CPI does not have a significant effect on the FTSE250 price over the 
long-run, lending support to the findings of 
\citeboth{gultekin1983} and \citeboth{firth1979}
that the Fisher hypothesis is valid for the UK. One of the 
reasons for this may be that, since stock prices represent the 
discounted value of future cash flows (which are real), 
they are not directly impacted by inflation in the long run. Moreover, 
firms are able to pass on nominal costs to consumers which reduces
the impact of inflation.
The freer role of the Bank of England in setting monetary policy could be 
another factor in this, as a stricter monetary policy would keep inflation
in check and obscure any long-term equilibrium relationship.

@@

The USD/GBP exchange rate significantly positively impacts the FTSE250 price, confirming the findings of 
\citeboth{wong2022}. That there is a significant relationship is not surprising, because whilst the FTSE250 is a more domestic indicator of the London Stock Exchange than the FTSE 100 Index, 
at times nearly $50\%$ of its earnings have come from abroad, indicating that 
the value of the Pound has a significant impact on this stock market. There are many reasons for why this effect is observed.
In the long run, the appreciation of the Pound would decrease the 
cost of imports, thereby increasing the 
profits of FTSE250 companies and inflating their share prices.
Moreover, a stronger Pound makes UK assets more appealing to foreign investors, as their investments would be worth more when converted back to their domestic currencies.
This would drive up demand for UK stocks and positively influence the 
FTSE250 price.

The interest rate is also found to have significant negative long-term and short-term effects on the stock market, confirming the findings of \citeboth{alam2009}, \citeboth{demir2019} and \citeboth{neifar2023}
This can be explained by the fact that higher interest 
rates increase the cost of borrowing for companies, leading to reduced investment 
and lower future earnings potential, in turn depressing stock prices. 
Furthermore, as per the flight-to-quality phenomenon, persistently higher interest rates make bonds and other fixed-income 
investments more attractive relative to stocks, leading investors to shift 
their portfolios away from equities and toward safer assets over time (\citeboth{asgharian2016}). 
In the short term, interest rate hikes immediately decrease the present value of earnings, as per the 
DCF valuation model, and depress prices.

@@

The model yields a mean absolute error of MAE = $0.07$ and 
$R^2 = 61\%$ for the short-term forecast. The results obtained are presented in Figure \ref{fig:shortmonthly}.
The significantly better performance of the model in the short-run
indicates that the selected macroeconomic variables possess
a stronger predictive power over shorter periods. This was highlighted 
earlier in the ARDL model, with the interest rate and money supply significantly
impacting the FTSE250 price in the short-run.

@@

\subsubsection{Long-term Forecast}
The model returns a mean absolute error of MAE $=0.16$ and 
$R^2 = 9\%$ for the long-term forecast, providing a 
$2.4\%$ increase in the $R^2$ obtained by the ARDL-ECM model. The results are presented 
in Figure \ref{fig:longmonthly}.

@@

Stock markets provide important information on the economy, 
and economic policymakers are vigilant in tracking the fluctuations 
in these markets, and try to forecast the 
forthcoming fluctuations based on policy changes that affect the macroeconomy.
Within this context, this study examined the 
impacts of the interest rate, USD/GBP
exchange rate, the Consumer Price Index, and 
M3 money supply market activity on
the FTSE250 over the October 1992 – March 2024 period.
The results suggested a cointegrated long-run
relationship between the macroeconomic variables and the stock market index. 
The long-run coefficients implied that
increases in the interest rate and depreciation in the exchange rate
significantly depress the FTSE250 value. 
Policymakers dealing with forecasting the UK stock markets
should consider the factors in the study, and to raise the stock
market's value, a lower interest rate regime should be adopted. This will 
stimulate consumer activity, reduce operating costs for firms, and make 
stocks more attractive investments, thereby increasing the stock market value. 
In addition, the Pound should be increased in value to lower import costs 
and increase the real value of firm earnings, both of which should drive up 
the value of the stock market. The combined effect of these changes will 
create a more stimulating business environment by increasing the capital 
available to firms for research and development, and more efficient 
capital and labour, in turn 
driving economic growth and prosperity. However, 
as with all economic policy shifts, these changes may have 
unintended consequences on other areas of the economy, potentially 
undermining any positive changes. Any policy changes must cautiously be made 
with this at the forefront.

@@

\newpage

\section*{Appendix}
\label{sec: appendix}

The \texttt{GaussianProcessRegressor} function in Python's \texttt{scikit-learn} library is based on the 
standard function
